Data Specialization for a Staged Lambda Calculus

### 2015-02-22 paper/splitting.tex
worked on examples and ground explanation

\subsection {Three World System?}

Why do we need a three world system?

>>>>

\subsection {Necessity of \bbonep}

We have not yet given a justification for why first stage code must be
partitioned between the worlds \bbonem\ and \bbonep.
Certainly one could imagine a simpler system without all the \texttt{grnd} annotations.  
Why wouldn't this work?

To see why, consider a function with a monostage body. 
For example, one which simply doubles its argument:
\begin{lstlisting}
2`prev{
  1`let fun double (n : int) = n + n
  in next{`2`hold {`1`double 5`2`}-7`1`}`2`
}`
\end{lstlisting}
Using the rule for splitting at \bbonem, this term splits into:
\begin{lstlisting}
1`let fun double (n : int) = (n+n, ())
in double 5`

2`l => 
let fun double (n : unit, l0 : unit) = ()
in (double (pi2 l); pi1 l)-7`
\end{lstlisting}

Observe that it's wasteful to run the stage \bbtwo\ version of \texttt{double}, since it always returns \texttt{()}. 
In this case, the wasted work is small, but if \texttt{double} were a recursive function,
then it's runtime may be unbounded.

Optimizing such an inefficiency away after-the-fact is in general a global program analysis,
since functions like \texttt{double} could be passed around as values.  
So instead of relying on hefty analysis, 
we take the approach of adding enough structure to the input language's type system to
allow the input code to {\em prove} itself to be---in some parts---monostage.

We find that giving input with \texttt{grnd} annotations is not especially cumbersome in practice.
Yet even if a language designer desired a different implicit/explicit trade-off,
then the three-world \lang\ would still be useful as a typed intermediate representation.
